---
title: "Error-in-variables regression for mother-egg concentrations"
output: html_document
---

In this script we will model the relationship between concentrations of organic
pollutants in paired samples of female (x-axis) and offspring (y-axis) tissues.
Since both female and egg tissues are dependent variables, we will consider an
error term on both these variables (following the idea of orthogonal
regression / error-in variables regression). 

Several models are developed and compared using wAIC. Model runs
are saved for post processing in a separate script.

Note: some of the code might take a while to execute, it might be better to run
the code chunk by chunk rather than knitting the whole document in one go.


```{r Setup, include=FALSE}

## script setup
## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# note: r version  4.3.2

# clear r environment
rm(list = ls())

# directories  ---> do not forget to set your working directory
dir.data <- "../Data/"
dir.modelruns <- "../Model runs/"
dir.jagsmodels <- "JAGS_models/"

# packages
library(dplyr)       # version 1.1.4
library(ggplot2)     # version 3.5.1
library(gridExtra)   # version 2.3
library(GGally)      # version 2.2.1
library(coda)        # version 0.19-4
library(rjags)       # version 4-15
library(R2jags)      # version 0.7-1
library(lattice)     # version 0.21-9
library(pracma)      # version 2.4.4 - for orthogonal regression using frequentist inference
library(loo)         # version 2.6.0

# for reproducibility
set.seed(124)


```



```{r ReadData, include=FALSE}

## read & prepare data
## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# read data
df_noncens <- read.csv(file = paste0(dir.data, "dataset_1.csv"))
dim(df_noncens)

# prepare data (i.e. log transform observations)
df_noncens <- df_noncens %>%  
  mutate(log10_Mother = log10(Mother_concentration)) %>%          
  mutate(log10_Offspring = log10(Offspring_concentration))
dim(df_noncens)  

# check for NA values
any(is.na(df_noncens$log10_Mother))
any(is.na(df_noncens$log10_Offspring))

```


```{r FreqModel, include=TRUE}

## frequentist orthogonal regression (on whole dataset)
## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

x_obs <- df_noncens$log10_Mother
y_obs <- df_noncens$log10_Offspring
odr <- odregress(x_obs, y_obs)    # Note: cannot have NA values in x or y, make sure to remove them
FreqCoeff <- odr$coeff
FreqCoeff

# Prediction
xnew <- seq(-2, 8, by = 0.5)
ynew <- cbind(xnew, 1) %*% odr$coeff

plot(x_obs, y_obs, xlim = c(-2, 8), ylim = c(-2, 8), las = 1)
abline(a = 0, b = 1)
abline(lm(y_obs ~ x_obs), col="blue")                               # normal linear regression line in blue
points(xnew, ynew, col = "red", pch = 17)
lines(c(-2, 8), FreqCoeff[1]*c(-2, 8) +  FreqCoeff[2], col="red")   # orthogonal regression in red

```

```{r Priors, include=FALSE}

## Priors
## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# take the largest concentration in either offspring or mother to inform on our
# prior distributions, then remove this largest point from the dataset (i.e. do
# not use the same data point twice)
# the max value is the multiplied by 3

max_xObs <- max(df_noncens$log10_Mother, na.rm=T)
max_yObs <- max(df_noncens$log10_Offspring, na.rm=T)

if(max_xObs > max_yObs){
  max <- max_xObs * 3
  df_noncens <- df_noncens[-which(df_noncens$log10_Mother == max_xObs), ]
}else{
  max <- max_yObs * 3
  df_noncens <- df_noncens[-which(df_noncens$log10_Offspring == max_yObs),]
}

```


```{r Cross-validation, include=TRUE}

## Cross-validation
## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Define sample size for k-folds over whole dataset (in our case 3 fold)
k <- 3
i <- nrow(df_noncens)
f <- as.integer(i/k) # size of a fraction

# Check if there are any data which could not be added to a fraction (i.e. the
# dataset size is not dividable by k) then, randomly select the k folds to add
# the remaining randomly selected sites
r <- i-(f*k)
i.fold <- rep(f, k)
if (r > 1){
  x <- sample(1:k, r)
  i.fold[x] <- i.fold[x] + 1
} else {
  x <- sample(1:k, 1)
  i.fold[x] <- i.fold[x] + r
}

# create vector where each fold integer is repeated i.fold times
# then, randomly sample without replacement
repeated_vector <- rep(1:k, times = i.fold)
fold.ind <- sample(repeated_vector, replace = F)
# hist(fold.ind)

# divide the data according to the random samples
fold1.df <- df_noncens[which(fold.ind==1),]
fold2.df <- df_noncens[which(fold.ind==2),]
fold3.df <- df_noncens[which(fold.ind==3),]

# Combine the folds into training and testing datasets
train1 <- rbind(fold1.df, fold2.df)
test1  <- fold3.df
train2 <- rbind(fold2.df, fold3.df)
test2  <- fold1.df
train3 <- rbind(fold1.df, fold3.df)
test3  <- fold2.df

# quick check
p1 <- ggplot(data = train1,
       aes(x = log10_Mother, y = log10_Offspring)) + geom_point()
p2 <- ggplot(data = train2,
       aes(x = log10_Mother, y = log10_Offspring)) + geom_point()
p3 <- ggplot(data = train3,
       aes(x = log10_Mother, y = log10_Offspring)) + geom_point()
grid.arrange(p1,p2,p3)

# need to save the datasets
datasets = list(training = list(CV1= train1,
                                CV2 = train2,
                                CV3 = train3),
                test = list(CV1= test1,
                            CV2 = test2,
                            CV3 = test3))
save(datasets,
     file = paste0(dir.modelruns, "datasets_", Sys.Date(),".RData"))


```



```{r CensoredData, include=FALSE}

## create censored data (from whole dataset, no CV)
## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# we will create datasets where 1 of the variables contains
# censored values (the lowest 20% of the data)

# calculate cut off
cutoff_Y <- quantile(df_noncens$log10_Offspring, probs = 0.2)
cutoff_X <- quantile(df_noncens$log10_Mother, probs = 0.2)

# Option 0: no censoring 
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# plot
png(file = "../Output/Uncensored data.png",
    res = 300, units = "cm", width = 12, height = 11)

p0 <- df_noncens %>% 
  ggplot() +
  geom_point(aes(x = log10_Mother, 
                 y = log10_Offspring)) + 
  scale_shape_manual(values=c(19, 4)) +
  labs(x = "Mother concentration",  
         y = "Offspring concentration") +
  theme_classic() + theme(legend.position = "none")
p0
dev.off()



# Option 1: censoring in Y variable (egg concentrations)
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# prepare data
df_cens1 <- df_noncens %>%  
  mutate(censored_Y_YN = ifelse(log10_Offspring >= cutoff_Y, 0, 1)) %>% # R variable
  mutate(LOD_Y = ifelse(log10_Offspring >= cutoff_Y, -10000, cutoff_Y)) %>% # cut off value, ridiculous number otherwise
  mutate(Yobs = ifelse(log10_Offspring >= cutoff_Y, log10_Offspring, NA)) # available offspring observations (NA for censored values)
# reorder the data with non censored followed by censored
data.obs <- df_cens1[which(df_cens1$censored_Y_YN == 0), ]     # subset: 411 
data.mis <- df_cens1[which(df_cens1$censored_Y_YN == 1), ]      # subset: 103
df_cens1 <- rbind(data.obs, data.mis) # total: 514 

# plot
png(file = "../Output/Censored Y data.png",
    res = 300, units = "cm", width = 12, height = 11)
p1 <- df_cens1 %>% 
  ggplot() +
  geom_point(aes(x = log10_Mother, 
                 y = log10_Offspring,
                 shape = as.factor(censored_Y_YN))) + 
  geom_hline(aes(yintercept = cutoff_Y),
              col = "#F26722", lwd = 1.2, lty = "dashed") +
  scale_shape_manual(values=c(19, 4)) +
  labs(x = "Mother concentration",  
       y = "Offspring concentration") +
  theme_classic() + theme(legend.position = "none")
p1
dev.off()

# Option 2: censoring in X variable (mother concentrations)
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# prepare data
df_cens2 <- df_noncens %>%  
  mutate(censored_X_YN = ifelse(log10_Mother >= cutoff_X, 0, 1)) %>% # R variable
  mutate(LOD_X = ifelse(log10_Mother >= cutoff_X, -10000, cutoff_X)) %>% # cut off value, ridiculous number otherwise
  mutate(Xobs = ifelse(log10_Mother >= cutoff_X, log10_Mother, NA)) # available offspring observations (NA for censored values)
# reorder the data with non censored followed by censored
data.obs <- df_cens2[which(df_cens2$censored_X_YN == 0), ]     # subset: 417 
data.mis <- df_cens2[which(df_cens2$censored_X_YN == 1), ]      # subset: 97
df_cens2 <- rbind(data.obs, data.mis) # total: 514 

# plot
png(file = "../Output/Censored X data.png",
    res = 300, units = "cm", width = 12, height = 11)
p2 <- df_cens2 %>% 
  ggplot() +
  geom_point(aes(x = log10_Mother, 
                 y = log10_Offspring,
                 shape = as.factor(censored_X_YN))) + 
  geom_vline(aes(xintercept = cutoff_X),
              col = "#F26722", lwd = 1.2, lty = "dashed") +
  scale_shape_manual(values=c(19, 4)) +
  labs(x = "Mother concentration",  
       y = "Offspring concentration") +
  theme_classic() + theme(legend.position = "none")
p2
dev.off()


# Option 3: censoring in X and Y variables (mother and offspring concentrations)
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# prepare data
df_cens3 <- df_noncens %>%  
  # censoring in Y
  mutate(censored_Y_YN = ifelse(log10_Offspring >= cutoff_Y, 0, 1)) %>% # 0 or 1 to indicate censoring
  mutate(LOD_Y = ifelse(log10_Offspring >= cutoff_Y, -10000, cutoff_Y)) %>% # value of the LOD
  mutate(Yobs = ifelse(log10_Offspring >= cutoff_Y, log10_Offspring, NA)) %>% # observed values
  # censoring in X
  mutate(censored_X_YN = ifelse(log10_Mother >= cutoff_X, 0, 1)) %>% # 0 or 1 to indicate censoring
  mutate(LOD_X = ifelse(log10_Mother >= cutoff_X, -10000, cutoff_X)) %>% # value of the LOD
  mutate(Xobs = ifelse(log10_Mother >= cutoff_X, log10_Mother, NA)) %>% # observed values
  # identify if censoring in both directions
  # mutate(censored_XandY = ifelse(censored_Y_YN + censored_X_YN == 2, 1, 0)) %>%
  mutate(cens_type = case_when(log10_Mother < cutoff_X & log10_Offspring < cutoff_Y ~ "both",
                            log10_Mother < cutoff_X & log10_Offspring >= cutoff_Y ~ "X only",
                            log10_Mother >= cutoff_X & log10_Offspring < cutoff_Y ~ "Y only",
                         log10_Mother >= cutoff_X & log10_Offspring >= cutoff_Y ~ "none"))
  

# reorder the data: (1) non censored followed by censored (2) in Y, (3) in X, and (4) in Y and X
data.obs <- df_cens3[which(df_cens3$cens_type == "none"), ] 
dim(data.obs)    # subset: 393
data.cen_Y <-  df_cens3[which(df_cens3$cens_type == "Y only"), ] 
dim(data.cen_Y) # subset: 24
data.cen_X <-  df_cens3[which(df_cens3$cens_type == "X only"), ] 
dim(data.cen_X) # subset: 18
data.cen_XY <-  df_cens3[which(df_cens3$cens_type == "both"), ] 
dim(data.cen_XY) # subset: 18
df_cens3 <- rbind(data.obs, data.cen_Y, data.cen_X, data.cen_XY) # total: 514 
dim(df_cens3) # total 514 rows

# plot
png(file = "../Output/Censored XY data.png",
    res = 300, units = "cm", width = 12, height = 11)
p3 <- df_cens3 %>% 
  ggplot() +
  geom_point(aes(x = log10_Mother, 
                 y = log10_Offspring,
                 shape = as.factor(cens_type))) + 
  geom_hline(aes(yintercept = cutoff_Y),
             col = "#F26722", lwd = 1.2, lty = "dashed") +
  geom_vline(aes(xintercept = cutoff_X),
             col = "#F26722", lwd = 1.2, lty = "dashed") +
  scale_shape_manual(values=c(4, 19, 4, 4)) +
  labs(x = "Mother concentration",  
       y = "Offspring concentration") +
  theme_classic() + theme(legend.position = "none")
p3
dev.off()

 
grid.arrange(p1,p2, p3, nrow=3)


# save datasets for later analysis and plotting
# need to save the datasets
datasets_censored = list(Ycens = df_cens1,
                         Xcens = df_cens2,
                         XYcens = df_cens3)
save(datasets_censored,
     file = paste0(dir.modelruns, "datasets_censored_", Sys.Date(),".RData"))
```

```{r Data, include=FALSE}

## Data
## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# data organised for JAGS - posterior sampling

# for the 1st analysis, using only uncensored data
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
data_M1_CV1 <- list(xobs = train1$log10_Mother,      # xobs = mother (see DAG for naming of variables and parameters)
                yobs = train1$log10_Offspring,       # yobs = Offspring
                n = nrow(train1),                    # number of observations 
                max = max)                           # to inform on prior
data_M1_CV2 <- list(xobs = train2$log10_Mother,    
                yobs = train2$log10_Offspring,  
                n = nrow(train2),            
                max = max)                  
data_M1_CV3 <- list(xobs = train3$log10_Mother,     
                yobs = train3$log10_Offspring,  
                n = nrow(train3),               
                max = max)  
              
data <- list(CV1 = data_M1_CV1,
             CV2 = data_M1_CV2,
             CV3 = data_M1_CV3)


# for the 2nd analysis, including censored data
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# censoring in Y
data_Cens_Y <- list(# Y values
                  yobs = df_cens1[which(df_cens1$censored_Y_YN == 0), "Yobs"],  # uncensored Yobs
                  z1   = df_cens1[which(df_cens1$censored_Y_YN == 1), "censored_Y_YN"], # type of censoring (1 = Left censored), for censored data
                  cut1  = df_cens1[which(df_cens1$censored_Y_YN == 1), "LOD_Y"],  # detection limit, for censored data
                  # X values
                  xobs = df_cens1$log10_Mother, # all X values
                  # nr. observations
                  n = nrow(df_cens1[which(df_cens1$censored_Y_YN == 0), ]), # length of non censored data
                  c1 = nrow(df_cens1[which(df_cens1$censored_Y_YN == 1), ]),  # length of censored data
                  # data for prior 
                  max = max) 

# Alternative, using the dinterval() function in JAGS
data_Cens_Y_dinterval <- list(
                  # Y values
                  yobs = df_cens1$Yobs,  
                  # X values
                  xobs = df_cens1$log10_Mother, 
                  # nr. observations
                  n = nrow(df_cens1[which(df_cens1$censored_Y_YN == 0), ]), # length of non censored data
                  J = nrow(df_cens1[which(df_cens1$censored_Y_YN == 1), ]),  # length of censored data
                  # censoring info 
                  R = rep(0, nrow(df_cens1[which(df_cens1$censored_Y_YN == 1), ])),     # left censoring R =0
                  lod = df_cens1[which(df_cens1$censored_Y_YN == 1), "LOD_Y"],
                  # data for prior 
                  max = max) 


# censoring in X
data_Cens_X <- list(# Y values
                  yobs = df_cens2$log10_Offspring,  # uncensored Yobs
                  # X values
                  xobs = df_cens2[which(df_cens2$censored_X_YN == 0), "Xobs"], # uncensored Xobs
                  z2   = df_cens2[which(df_cens2$censored_X_YN == 1), "censored_X_YN"], # type of censoring (1 = Left censored), for censored data
                  cut2 = df_cens2[which(df_cens2$censored_X_YN == 1), "LOD_X"],  # detection limit, for censored data
                  # nr. observations
                  n = nrow(df_cens2[which(df_cens2$censored_X_YN == 0), ]), # length of non censored data
                  c2 = nrow(df_cens2[which(df_cens2$censored_X_YN == 1), ]),  # length of censored data in X direction
                  # data for prior 
                  max = max) 


# censoring in Y and X
data_Cens_YandX <- list(
  n = nrow(df_cens3[which(df_cens3$cens_type == "none"), ]), # length of non-censored data
  c1 = nrow(df_cens3[which(df_cens3$cens_type == "Y only"), ]), # length of data censored in Y
  c2 = nrow(df_cens3[which(df_cens3$cens_type == "X only"), ]), # length of data censored in X
  c3 = nrow(df_cens3[which(df_cens3$cens_type == "both"), ]), # length of completely censored data
  # observations - continuous variable
  yobs = df_cens3[which(df_cens3$cens_type != "both"), "Yobs"],  # uncensored Yobs
  xobs = df_cens3[which(df_cens3$cens_type != "both"), "Xobs"], # uncensored Xobs
  # observations - binary variable
  z1 = df_cens3[which(df_cens3$cens_type ==  "Y only"), "censored_Y_YN"], # type of censoring (1 = Left censored), for censored data
  z2 = df_cens3[which(df_cens3$cens_type ==  "X only"), "censored_X_YN"], 
  z3 = df_cens3[which(df_cens3$cens_type ==  "both"), "censored_Y_YN"], 
  z4 = df_cens3[which(df_cens3$cens_type ==  "both"), "censored_X_YN"], 
  # detection limits (i.e. censoring cutoffs)
  cut1 = df_cens3[which(df_cens3$cens_type ==  "Y only"), "LOD_Y"],  # detection limit, for censored data
  cut2 = df_cens3[which(df_cens3$cens_type ==  "X only"), "LOD_X"],
  cut3 = df_cens3[which(df_cens3$cens_type ==  "both"), "LOD_Y"],  
  cut4 = df_cens3[which(df_cens3$cens_type ==  "both"), "LOD_X"],
  # data for prior
  max = max)


```


```{r ReadModel, include=FALSE}

## Define model
## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# models for uncensored data ~~~~~~~~~~~~~~~
# full, 6 parameter model
modelM1a <- "M1a.txt"
parametersM1a <- c("a", "b", "mu_x", "tau_x", "tau_y", "tau_u") # mu_x and tau_x are the population level params

# reduced, 5 parameter model (equal variance on Xobsi and Xi)
modelM1b <- "M1b.txt"
parametersM1b <- c("a", "b", "mu_x", "tau_x", "tau")

# reduced, 4 parameter model (fixed intercept)
modelM1c <- "M1c.txt"
parametersM1c <- c("a", "mu_x", "tau_x", "tau")


# model incl. censored data ~~~~~~~~~~~~~~~
# reduced, 5 parameter model for censored data
modelM1b_cens_Y <- "M1b - censored_Y.txt" # cens in Y
modelM1b_cens_Yalt <- "M1b - censored_Y_dinterval.txt" # Alternative: using dinterval() function in JAGS
parametersM1b <- c("a", "b", "mu_x", "tau_x", "tau")
modelM1b_cens_X <- "M1b - censored_X.txt" # cens in X
modelM1b_cens_YandX <- "M1b - censored_XY.txt" # cens in X
parametersM1b <- c("a", "b", "mu_x", "tau_x", "tau")

```


```{r Trial, include=FALSE}

## Run trail model to check how long chains should be (tested on 1 CV set)
## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# settings
n_chains   <- 3


# uncensored models 
# -----------------------------------------------------------
# -----------------------------------------------------------

# full, 6 parameter model
# -----------------------------------------------------------
JagsM1a_trial <- jags(model.file = paste0(dir.jagsmodels, modelM1a),
               parameters.to.save = parametersM1a,
               data = data_M1_CV1,
               # inits = initsList,
               n.chains = n_chains,
               n.burnin = 5000,
               n.iter = 20000,
               n.thin = 10)
ChainsM1a_trial <- as.mcmc(JagsM1a_trial)
PosteriorM1a_trial <- as.data.frame(as.matrix(ChainsM1a_trial))
# Raftery and Lewis's diagnostic (check needed length of MCMC samples)
RLD_M1a <- raftery.diag(PosteriorM1a_trial)
RLD_M1a$resmatrix
M1a_thin <- round(max(RLD_M1a$resmatrix[, "I"]) + 0.5)
M1a_niter <- max(RLD_M1a$resmatrix[, "Nmin"]) * M1a_thin
M1a_burnin <- max(RLD_M1a$resmatrix[, "M"])



# reduced, 5 parameter model (equal variance on Xobsi and Xi)
# -----------------------------------------------------------
JagsM1b_trial <- jags(model.file = paste0(dir.jagsmodels, modelM1b),
               parameters.to.save = parametersM1b,
               data = data_M1_CV1,
               # inits = initsList,
               n.chains = n_chains,
               n.burnin = 5000,
               n.iter = 20000,
               n.thin = 10)
ChainsM1b_trial <- as.mcmc(JagsM1b_trial)
PosteriorM1b_trial <- as.data.frame(as.matrix(ChainsM1b_trial))
# Raftery and Lewis's diagnostic (check needed length of MCMC samples)
RLD_M1b <- raftery.diag(PosteriorM1b_trial)
RLD_M1b$resmatrix
M1b_thin <- round(max(RLD_M1b$resmatrix[, "I"]) + 0.5)
M1b_niter <- max(RLD_M1b$resmatrix[, "Nmin"]) * M1b_thin
M1b_burnin <- max(RLD_M1b$resmatrix[, "M"])


# reduced, 4 parameter model (fixed intercept)
# -----------------------------------------------------------
JagsM1c_trial <- jags(model.file = paste0(dir.jagsmodels, modelM1c),
               parameters.to.save = parametersM1c,
               data = data_M1_CV1,
               # inits = initsList,
               n.chains = n_chains,
               n.burnin = 5000,
               n.iter = 20000,
               n.thin = 10)
ChainsM1c_trial <- as.mcmc(JagsM1c_trial)
PosteriorM1c_trial <- as.data.frame(as.matrix(ChainsM1c_trial))
# Raftery and Lewis's diagnostic (check needed length of MCMC samples)
RLD_M1c <- raftery.diag(PosteriorM1c_trial)
RLD_M1c$resmatrix
M1c_thin <- round(max(RLD_M1c$resmatrix[, "I"]) + 0.5)
M1c_niter <- max(RLD_M1c$resmatrix[, "Nmin"]) * M1c_thin
M1c_burnin <- max(RLD_M1c$resmatrix[, "M"])


# censored models 
# -----------------------------------------------------------
# -----------------------------------------------------------

# M5 param, censored in Y
# -----------------------------------------------------------
JagsM1b_cens_Y_trial <- jags(model.file = paste0(dir.jagsmodels, modelM1b_cens_Y),
               parameters.to.save = parametersM1b,
               data = data_Cens_Y,
               # inits = initsList,
               n.chains = n_chains,
               n.burnin = 5000,
               n.iter = 20000,
               n.thin = 10)
ChainsM1b_cens_Y_trial <- as.mcmc(JagsM1b_cens_Y_trial)
PosteriorM1b_cens_Y_trial <- as.data.frame(as.matrix(ChainsM1b_cens_Y_trial))
# Raftery and Lewis's diagnostic (check needed length of MCMC samples)
RLD_M1b_Y <- raftery.diag(PosteriorM1b_cens_Y_trial)
RLD_M1b_Y$resmatrix
M1b_cens_Y_thin <- round(max(RLD_M1b_Y$resmatrix[, "I"]) + 0.5)
M1b_cens_Y_niter <- max(RLD_M1b_Y$resmatrix[, "Nmin"]) * M1b_cens_Y_thin
M1b_cens_Y_burnin <- max(RLD_M1b_Y$resmatrix[, "M"])

# Alternative using dinterval() function in JAGS
JagsM1b_cens_Y_trialalt <- jags(model.file = paste0(dir.jagsmodels, modelM1b_cens_Yalt),
               parameters.to.save = parametersM1b,
               data = data_Cens_Y_dinterval,
               # inits = initsList,
               n.chains = n_chains,
               n.burnin = 5000,
               n.iter = 20000,
               n.thin = 10)
ChainsM1b_cens_Y_trialalt <- as.mcmc(JagsM1b_cens_Y_trialalt)
PosteriorM1b_cens_Y_trialalt <- as.data.frame(as.matrix(ChainsM1b_cens_Y_trialalt))
#### Note: this implementation is not further pursued, it was just presented here
# as some users (reviewers) might use the dinterva approach. Clearly, this also works
# we chose to align closey with the mathematical implementation presented in eqn. 2
# of the manuscript, where the bernoulli distribution is explicitely used (hence, 
# this alternative with "dinterval()" is not worked out bejond this point)




# M5 param, censored in X
# -----------------------------------------------------------
JagsM1b_cens_X_trial <- jags(model.file = paste0(dir.jagsmodels, modelM1b_cens_X),
               parameters.to.save = parametersM1b,
               data = data_Cens_X,
               # inits = initsList,
               n.chains = n_chains,
               n.burnin = 5000,
               n.iter = 20000,
               n.thin = 10)
ChainsM1b_cens_X_trial <- as.mcmc(JagsM1b_cens_X_trial)
PosteriorM1b_cens_X_trial <- as.data.frame(as.matrix(ChainsM1b_cens_X_trial))
# Raftery and Lewis's diagnostic (check needed length of MCMC samples)
RLD_M1b_X <- raftery.diag(PosteriorM1b_cens_X_trial)
RLD_M1b_X$resmatrix
M1b_cens_X_thin <- round(max(RLD_M1b_X$resmatrix[, "I"]) + 0.5)
M1b_cens_X_niter <- max(RLD_M1b_X$resmatrix[, "Nmin"]) * M1b_cens_X_thin
M1b_cens_X_burnin <- max(RLD_M1b_X$resmatrix[, "M"])


# M5 param, censored in X and Y
# -----------------------------------------------------------
JagsM1b_cens_XY_trial <- jags(model.file = paste0(dir.jagsmodels, modelM1b_cens_YandX),
               parameters.to.save = parametersM1b,
               data = data_Cens_YandX,
               # inits = initsList,
               n.chains = n_chains,
               n.burnin = 5000,
               n.iter = 20000,
               n.thin = 10)
ChainsM1b_cens_XY_trial <- as.mcmc(JagsM1b_cens_XY_trial)
PosteriorM1b_cens_XY_trial <- as.data.frame(as.matrix(ChainsM1b_cens_XY_trial))
# Raftery and Lewis's diagnostic (check needed length of MCMC samples)
RLD_M1b_XY <- raftery.diag(PosteriorM1b_cens_XY_trial)
RLD_M1b_XY$resmatrix
M1b_cens_XY_thin <- round(max(RLD_M1b_XY$resmatrix[, "I"]) + 0.5)
M1b_cens_XY_niter <- max(RLD_M1b_XY$resmatrix[, "Nmin"]) * M1b_cens_XY_thin
M1b_cens_XY_burnin <- max(RLD_M1b_XY$resmatrix[, "M"])


```



```{r PosteriorModel}

# Note: this part might take some time

## Run model with data -> i.e. sample the posterior distribution
## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

n_chains <- 3


# uncensored models 
# -----------------------------------------------------------
# -----------------------------------------------------------

# full, 6 parameter model
# -----------------------------------------------------------

start <- Sys.time()

JagsM1a <- list()
for(i in 1:length(data)){
  JagsM1a[[i]] <- jags(model.file = paste0(dir.jagsmodels, modelM1a),
                     parameters.to.save = parametersM1a,
                     data = data[[i]],
                     # inits = initsList,
                     n.chains = n_chains,
                     n.burnin = max(100, round(M1a_burnin, -1) *3),  
                     n.iter = round(M1a_niter, -1) *3,   
                     n.thin = 10)
}

time.elapsed <- Sys.time() - start
time.elapsed

save(JagsM1a,
      file = paste0(dir.modelruns, "runs_M1a_", Sys.Date(),".RData"))



# reduced, 5 parameter model (equal variance on Xobsi and Xi)
# -----------------------------------------------------------

start <- Sys.time()

JagsM1b <- list()
for(i in 1:length(data)){
  JagsM1b[[i]] <- jags(model.file = paste0(dir.jagsmodels, modelM1b),
                     parameters.to.save = parametersM1b,
                     data = data[[i]],
                     # inits = initsList,
                     n.chains = n_chains,
                     n.burnin = max(100, round(M1b_burnin, -1) *3),  
                     n.iter = round(M1b_niter, -1) *3,   
                     n.thin = 10)
}

time.elapsed <- Sys.time() - start
time.elapsed

save(JagsM1b,
      file = paste0(dir.modelruns, "runs_M1b_", Sys.Date(),".RData"))


# reduced, 4 parameter model (fixed intercept)
# -----------------------------------------------------------

start <- Sys.time()

JagsM1c <- list()
for(i in 1:length(data)){
  JagsM1c[[i]] <- jags(model.file = paste0(dir.jagsmodels, modelM1c),
                     parameters.to.save = parametersM1c,
                     data = data[[i]],
                     # inits = initsList,
                     n.chains = n_chains,
                     n.burnin = max(100, round(M1c_burnin, -1) *3),  
                     n.iter = round(M1c_niter, -1) *3,   
                     n.thin = 10)
}

time.elapsed <- Sys.time() - start
time.elapsed

save(JagsM1c,
      file = paste0(dir.modelruns, "runs_M1c_", Sys.Date(),".RData"))


# censored models 
# -----------------------------------------------------------
# -----------------------------------------------------------

# M5 param, censored in Y
# -----------------------------------------------------------
start <- Sys.time()
JagsM1b_cens_Y <- jags(model.file = paste0(dir.jagsmodels, modelM1b_cens_Y),
                       parameters.to.save = parametersM1b,
                       data = data_Cens_Y,
                       # inits = initsList,
                       n.chains = n_chains,
                       n.burnin = max(100, round(M1b_cens_Y_burnin, -1) *3),  
                       n.iter = round(M1b_cens_Y_niter, -1) *3,   
                       n.thin = 10)
time.elapsed <- Sys.time() - start
time.elapsed

save(JagsM1b_cens_Y,
      file = paste0(dir.modelruns, "runs_M1b_cens_Y_", Sys.Date(),".RData"))


# M5 param, censored in X
# -----------------------------------------------------------
start <- Sys.time()
JagsM1b_cens_X <- jags(model.file = paste0(dir.jagsmodels, modelM1b_cens_X),
                       parameters.to.save = parametersM1b,
                       data = data_Cens_X,
                       # inits = initsList,
                       n.chains = n_chains,
                       n.burnin = max(100, round(M1b_cens_X_burnin, -1) *3),  
                       n.iter = round(M1b_cens_X_niter, -1) *3,   
                       n.thin = 10)
time.elapsed <- Sys.time() - start
time.elapsed

save(JagsM1b_cens_X,
      file = paste0(dir.modelruns, "runs_M1b_cens_X_", Sys.Date(),".RData"))

# M5 param, censored in Y and X
# -----------------------------------------------------------

start <- Sys.time()
JagsM1b_cens_XY <- jags(model.file = paste0(dir.jagsmodels, modelM1b_cens_YandX),
                       parameters.to.save = parametersM1b,
                       data = data_Cens_YandX,
                       # inits = initsList,
                       n.chains = n_chains,
                       n.burnin = max(100, round(M1b_cens_XY_burnin, -1) *3),  
                       n.iter = round(M1b_cens_XY_niter, -1) *3,   
                       n.thin = 10)
time.elapsed <- Sys.time() - start
time.elapsed

save(JagsM1b_cens_XY,
      file = paste0(dir.modelruns, "runs_M1b_cens_XY_", Sys.Date(),".RData"))


```

```{r CompareModels, echo = TRUE}

## Compare models: wAIC 
## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# wAIC = Watanabe-Akaike information criterion aka. widely applicable
# information criterion the model with lower wAIC is preferred
# based on example from Olivier Gimenez 

# uncensored models 
# -----------------------------------------------------------
# -----------------------------------------------------------
# full, 6 parameter model
# ----------------------------------------------------------

# get WAIC
summarysM1a <- list()
for(i in 1:length(data)){
  samples <- jags.samples(JagsM1a[[i]]$model,
                                  c("WAIC","deviance"),
                                  type = "mean",
                                  n.iter = 5000,
                                  n.burnin = 1000,
                                  n.thin = 1)
  samples$p_waic <- samples$WAIC
  samples$waic <- samples$deviance + samples$p_waic
  summarysM1a[[i]] <- sapply(samples, sum)
}

summarysM1a
waic_M1a <- c(summarysM1a[[1]][["waic"]], summarysM1a[[2]][["waic"]], summarysM1a[[3]][["waic"]])
cat("wAIC Mean: ", mean(waic_M1a), "Sd: ", sd(waic_M1a))
pwaic_M1a <- c(summarysM1a[[1]][["p_waic"]], summarysM1a[[2]][["p_waic"]], summarysM1a[[3]][["p_waic"]])
cat("p_wAIC Mean: ", mean(pwaic_M1a), "Sd: ", sd(pwaic_M1a))
deviance_M1a <- c(summarysM1a[[1]][["deviance"]], summarysM1a[[2]][["deviance"]], summarysM1a[[3]][["deviance"]])
cat("deviance Mean: ", mean(deviance_M1a), "Sd: ", sd(deviance_M1a))


# gelman diagnostic
gd1 <- gelman.diag(as.mcmc(JagsM1a[[1]]))
gd1 <- gd1$psrf[c("a", "mu_x", "tau_x", "tau_y", "b", "tau_u"),]
gd2 <- gelman.diag(as.mcmc(JagsM1a[[2]]))
gd2 <- gd2$psrf[c("a", "mu_x", "tau_x", "tau_y", "b", "tau_u"),]
gd3 <- gelman.diag(as.mcmc(JagsM1a[[3]]))
gd3 <- gd3$psrf[c("a", "mu_x", "tau_x", "tau_y", "b", "tau_u"),]

res <- data.frame(CV1 = gd1[,"Point est."],
                  CV2 = gd2[,"Point est."],
                  CV3 = gd3[,"Point est."])
apply(res, 1, mean)

# reduced, 5 parameter model (equal variance on Xobsi and Xi)
# ----------------------------------------------------------


# get WAIC
summarysM1b <- list()
for(i in 1:length(data)){
  samples <- jags.samples(JagsM1b[[i]]$model,
                                  c("WAIC","deviance"),
                                  type = "mean",
                                  n.iter = 5000,
                                  n.burnin = 1000,
                                  n.thin = 1)
  samples$p_waic <- samples$WAIC
  samples$waic <- samples$deviance + samples$p_waic
  summarysM1b[[i]] <- sapply(samples, sum)
}

summarysM1b
waic_M1b <- c(summarysM1b[[1]][["waic"]], summarysM1b[[2]][["waic"]], summarysM1b[[3]][["waic"]])
cat("wAIC Mean: ", mean(waic_M1b), "Sd: ", sd(waic_M1b))
pwaic_M1b <- c(summarysM1b[[1]][["p_waic"]], summarysM1b[[2]][["p_waic"]], summarysM1b[[3]][["p_waic"]])
cat("p_wAIC Mean: ", mean(pwaic_M1b), "Sd: ", sd(pwaic_M1b))
deviance_M1b <- c(summarysM1b[[1]][["deviance"]], summarysM1b[[2]][["deviance"]], summarysM1b[[3]][["deviance"]])
cat("deviance Mean: ", mean(deviance_M1b), "Sd: ", sd(deviance_M1b))

# gelman diagnostic
gd1 <- gelman.diag(as.mcmc(JagsM1b[[1]]))
gd1 <- gd1$psrf[c("a", "mu_x", "tau_x", "tau", "b"),]
gd2 <- gelman.diag(as.mcmc(JagsM1b[[2]]))
gd2 <- gd2$psrf[c("a", "mu_x", "tau_x", "tau", "b"),]
gd3 <- gelman.diag(as.mcmc(JagsM1b[[3]]))
gd3 <- gd3$psrf[c("a", "mu_x", "tau_x", "tau", "b"),]

res <- data.frame(CV1 = gd1[,"Point est."],
                  CV2 = gd2[,"Point est."],
                  CV3 = gd3[,"Point est."])
apply(res, 1, mean)

# reduced, 4 parameter model (fixed intercept)
# ----------------------------------------------

# get WAIC
summarysM1c <- list()
for(i in 1:length(data)){
  samples <- jags.samples(JagsM1c[[i]]$model,
                                  c("WAIC","deviance"),
                                  type = "mean",
                                  n.iter = 5000,
                                  n.burnin = 1000,
                                  n.thin = 1)
  samples$p_waic <- samples$WAIC
  samples$waic <- samples$deviance + samples$p_waic
  summarysM1c[[i]] <- sapply(samples, sum)
}

summarysM1c
waic_M1c <- c(summarysM1c[[1]][["waic"]], summarysM1c[[2]][["waic"]], summarysM1c[[3]][["waic"]])
cat("wAIC Mean: ", mean(waic_M1c), "Sd: ", sd(waic_M1c))
pwaic_M1c <- c(summarysM1c[[1]][["p_waic"]], summarysM1c[[2]][["p_waic"]], summarysM1c[[3]][["p_waic"]])
cat("p_wAIC Mean: ", mean(pwaic_M1c), "Sd: ", sd(pwaic_M1c))
deviance_M1c <- c(summarysM1c[[1]][["deviance"]], summarysM1c[[2]][["deviance"]], summarysM1c[[3]][["deviance"]])
cat("deviance Mean: ", mean(deviance_M1c), "Sd: ", sd(deviance_M1c))

# gelman diagnostic
gd1 <- gelman.diag(as.mcmc(JagsM1c[[1]]))
gd1 <- gd1$psrf[c("a", "mu_x", "tau_x", "tau"),]
gd2 <- gelman.diag(as.mcmc(JagsM1c[[2]]))
gd2 <- gd2$psrf[c("a", "mu_x", "tau_x", "tau"),]
gd3 <- gelman.diag(as.mcmc(JagsM1c[[3]]))
gd3 <- gd3$psrf[c("a", "mu_x", "tau_x", "tau"),]

res <- data.frame(CV1 = gd1[,"Point est."],
                  CV2 = gd2[,"Point est."],
                  CV3 = gd3[,"Point est."])
apply(res, 1, mean)


# censored models 
# -----------------------------------------------------------
# -----------------------------------------------------------

# reduced, 5 parameter model  incl censored data
# -----------------------------------------------------------

# M5 param, censored in Y
# -----------------------------------------------------------
# gelman diagnostic
gd <- gelman.diag(as.mcmc(JagsM1b_cens_Y))
gd <- gd$psrf[c("a", "mu_x", "tau_x", "tau", "b"),]
gd

# M5 param, censored in X
# -----------------------------------------------------------
# gelman diagnostic
gd <- gelman.diag(as.mcmc(JagsM1b_cens_X))
gd <- gd$psrf[c("a", "mu_x", "tau_x", "tau", "b"),]
gd

# M5 param, censored in Y and X
# -----------------------------------------------------------
# gelman diagnostic
gd <- gelman.diag(as.mcmc(JagsM1b_cens_XY))
gd <- gd$psrf[c("a", "mu_x", "tau_x", "tau", "b"),]
gd



```



